---
title: "Sound_hab_nmds"
author: "Becca Van Hoeck"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(vegan)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Vegan NMDS - pmHT site average

```{r}
Df_pmHT = read.csv("data/clean_data/Df_pmHT_1_square.csv", header = F) 
Df_pmHT = as.dist(Df_pmHT)
Df_pmHT_metadata = read.csv("data/metadata/Df_pmHT_1_metadata.csv", header = T)

nmds_1 = metaMDS(Df_pmHT, k=3, trymax = 20, autotransform = FALSE, plot = FALSE)

nmdsScores = as.data.frame(scores(nmds_1))  
nmdsScores = cbind(nmdsScores,Df_pmHT_metadata)
nmdsScores$deploy = factor(nmdsScores$deploy)

# habitat: 
ggplot(aes(x = NMDS1, y = NMDS2, z = NMDS3, color = habitat), data = nmdsScores)+ theme_bw()+
  geom_point(size = 3)+
  geom_text(aes(label = site), hjust = 1.5)

# Deploy alone: mostly buckshot, suggests there are deeper patterns than seasonal 
ggplot(aes(x = NMDS1, y = NMDS2), data = nmdsScores)+ theme_bw()+
  geom_point(size = 3, aes(color = deploy))

# habitat and deploy
ggplot(aes(x = NMDS1, y = NMDS2), data = nmdsScores)+ theme_bw()+
  geom_point(size = 3, aes(color = habitat, shape = deploy))


```

## Add Environmental Vector

```{r}
env_data = read.csv("data/clean_data/env_metrics.csv", header = TRUE)
env_cor = cor(env_data[,c(4:23)], use = "pairwise.complete.obs")
#write.csv(env_cor,"data/clean_data/nmds_env_cor.csv")
#without seagrass only metrics
env_filt = env_data %>%
  select(-index, -deploy_num.x, -site, -bare, -evenness, -seagrass_comb, -shannonD_comb, -evenness_comb,
         -totalcover, -avg_shoot_density, -avg_blade_length, -max_10, -max_25perc)

# need to assess normality and consider transformations
# convert to z-scores?
env_zscores = as.data.frame(sapply(env_filt, function(env_filt) (env_filt-mean(env_filt))/sd(env_filt)))


#test with vegan envfit
#env_test= envfit(nmds_1, env_filt, perm = 999, na.rm = TRUE)
env_test = envfit(nmds_1 ~ halodule + zostera + ruppia + algae + oyster + saltmarsh + other + 
                     richness + shannonD + richness_comb, data = env_zscores, perm = 999)
env_test
envScores = as.data.frame(scores(env_test, "vectors"))

# habitat and deploy
ggplot(aes(x = NMDS1, y = NMDS2), data = nmdsScores)+ theme_bw()+
  geom_point(size = 3, aes(color = habitat, shape = deploy))+
  geom_text(aes(label = site), hjust = 1.5)+
  geom_segment(data =envScores ,aes(x=0,xend=NMDS1,y=0,yend=NMDS2),
               arrow = arrow(length = unit(0.5, "cm")),colour="black")+ 
  geom_text(data=envScores,aes(x=NMDS1,y=NMDS2,label=rownames(envScores)),size=4)+
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank())+
  geom_hline(yintercept = 0, color = "grey")+
  geom_vline(xintercept = 0, color = "grey")

# nmds with only significant env vectors
envScoresSig = envScores[c(1,5),]
ggplot(aes(x = NMDS1, y = NMDS2), data = nmdsScores)+ theme_bw()+
  geom_point(size = 3, aes(color = habitat, shape = deploy))+
  geom_text(aes(label = site), hjust = 1.5)+
  geom_segment(data =envScoresSig ,aes(x=0,xend=NMDS1,y=0,yend=NMDS2),
               arrow = arrow(length = unit(0.5, "cm")),colour="black")+ 
  geom_text(data=envScoresSig,aes(x=NMDS1,y=NMDS2,label=rownames(envScoresSig)),size=4)+
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank())+
  geom_hline(yintercept = 0, color = "grey")+
  geom_vline(xintercept = 0, color = "grey")

# only halodule was significant with permanova
permanova_test = adonis(Df_pmHT ~ halodule + zostera + ruppia + algae + oyster + saltmarsh + other + 
                          richness + shannonD + richness_comb, data = env_zscores, perm = 999) #, 
                        #stata = Df_pmHT_metadata$habitat)
permanova_test

# Permdisp - evaluating differences between habitat types
permdisp_test = betadisper(Df_pmHT, group = Df_pmHT_metadata$habitat)
permdisp_test

TukeyHSD(permdisp_test, conf.level = 0.95)


# ordellipse (vegan package)
#hab_ellipse <-ordiellipse(pmHT_nms, pmHT_nms_df$habitat, display = "sites", 
#                   kind = "se", conf = 0.95, label = T)
```

## NMDS site average - within habitat variation
- Seagrass: nothing immediately interesting at first glance. No real grouping. 
            Investigate spectra to decide if it's worth pursuing further
- Creek: no convergence. Not enough data. 
- Mudflat: didn't even try. 

- Could group creek and mudflat together. Then it might have enough data
    - only converged with k = 2, with 50 solution attempts.
    - many points plotted directly on top of one another. Are they really that similar? 

```{r}
#Seagrass
Df_pmHT_seagrass = read.csv("data/clean_data/Df_pmHT_1_square_seagrass.csv", header = F) 
Df_pmHT_seagrass = as.dist(Df_pmHT_seagrass)
Df_pmHT_metadata = read.csv("data/metadata/Df_pmHT_1_metadata.csv", header = T)

seagrass_metadata = Df_pmHT_metadata[c(2,3,6,7,10,11,14,15,19,20,23,24),]
seagrass_env = env_data[c(2,3,6,7,10,11,14,15,19,20,23,24),]
seagrass_env_cor = cor(seagrass_env[,c(4:23)], use = "pairwise.complete.obs")
#write.csv(seagrass_env_cor,"data/clean_data/nmds_env_cor_seagrass.csv")
env_filt_seagrass = seagrass_env %>%
  select(-index, -deploy_num.x, -site, -bare, -oyster, -saltmarsh, -evenness, -seagrass_comb,  -evenness_comb,
         -shannonD_comb, -max_25perc)

nmds_1_seagrass = metaMDS(Df_pmHT_seagrass, k=3, trymax = 20, autotransform = FALSE, plot = FALSE)
nmdsScores_seagrass = as.data.frame(scores(nmds_1_seagrass))  
nmdsScores_seagrass = cbind(nmdsScores_seagrass,seagrass_metadata)
nmdsScores_seagrass$deploy = factor(nmdsScores_seagrass$deploy)

env_test_seagrass= envfit(nmds_1_seagrass, env_filt_seagrass, perm = 999)
envScores_seagrass = as.data.frame(scores(env_test_seagrass, "vectors"))

ggplot(aes(x = NMDS1, y = NMDS2), data = nmdsScores_seagrass)+ theme_bw()+
  geom_point(size = 3, aes(color = deploy))+
  geom_text(aes(label = site), hjust = 1.5)+
  geom_segment(data =envScores_seagrass ,aes(x=0,xend=NMDS1,y=0,yend=NMDS2),
               arrow = arrow(length = unit(0.5, "cm")),colour="black")+ 
  geom_text(data=envScores_seagrass,aes(x=NMDS1,y=NMDS2,label=rownames(envScores_seagrass)),size=4)+
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank())+
  geom_hline(yintercept = 0, color = "grey")+
  geom_vline(xintercept = 0, color = "grey")

# Creek - no convergence - insufficient data
# added mudflat to creek since they were all adjacent to saltmarsh
Df_pmHT_creek = read.csv("data/clean_data/Df_pmHT_1_square_creek.csv", header = F) 
Df_pmHT_creek = as.dist(Df_pmHT_creek)
Df_pmHT_metadata = read.csv("data/metadata/Df_pmHT_1_metadata.csv", header = T)

creek_metadata = Df_pmHT_metadata[c(1,4,5,8,9,12,13,16,17,18,21,22),]
creek_env = env_data[c(1,4,5,8,9,12,13,16,17,18,21,22),]
creek_env_cor = cor(creek_env[,c(4:23)], use = "pairwise.complete.obs")
#write.csv(creek_env_cor,"data/clean_data/nmds_env_cor_creek_mudflat.csv")
env_filt_creek = creek_env %>%
  select(-index, -deploy_num.x, -site, -bare, -zostera, -ruppia, -algae, shannonD, -evenness, -evenness_comb,
         -totalcover, -avg_shoot_density, -avg_blade_length, -max_10, -max_25perc)

nmds_1_creek = metaMDS(Df_pmHT_creek, k=2, trymax = 50, autotransform = FALSE, plot = FALSE)
nmdsScores_creek = as.data.frame(scores(nmds_1_creek))  
nmdsScores_creek = cbind(nmdsScores_creek,creek_metadata)
nmdsScores_creek$deploy = factor(nmdsScores_creek$deploy)

env_test_creek= envfit(nmds_1_creek, env_filt_creek, perm = 999)
envScores_creek = as.data.frame(scores(env_test_creek, "vectors"))

ggplot(aes(x = NMDS1, y = NMDS2), data = nmdsScores_creek)+ theme_bw()+
  geom_point(size = 3, aes(color = deploy, shape = habitat))+
  geom_text(aes(label = site), hjust = 1.5)+
  geom_segment(data =envScores_creek ,aes(x=0,xend=NMDS1,y=0,yend=NMDS2),
               arrow = arrow(length = unit(0.5, "cm")),colour="black")+ 
  geom_text(data=envScores_creek,aes(x=NMDS1,y=NMDS2,label=rownames(envScores_creek)),size=4)+
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank())+
  geom_hline(yintercept = 0, color = "grey")+
  geom_vline(xintercept = 0, color = "grey")

```



