---
title: "Sound_hab_spl_correlations"
author: "Becca Van Hoeck"
date: "3/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and data
```{r}
library(ggplot2)
library(GGally)
library(dplyr)
library(tidyr)

ecometrics = read.csv("data/clean_data/ecometrics.csv")
ACI = read.csv("data/clean_data/avg_ACI1024.csv")
H_entropy = read.csv("data/clean_data/avg_H1024.csv")

ecometrics$ACI = ACI$avg_ACI1024
ecometrics$entropy= H_entropy$avg_H1024

# make deployment a factor to include as a random effect
ecometrics$deploy = factor(ecometrics$deploy_num.x)
```

## Explore correlations:
1. Examine correlation within habitat/community metrics and remove redundant variables (corr > 0.7)
2. Explore correlation between habitat/community metrics and Sound Pressure Level

Start with habitat then community to reduce the number of correlations viewing at a single time
```{r}
# HABITAT
sound_hab = ecometrics %>%
  select(site, habitat, totalcover, seagrass_comb, oyster, saltmarsh, richness, evenness,
         avg_shoot_density, avg_blade_length,
        avg_spl_low, avg_spl_high, ACI, entropy)

ggpairs(sound_hab[,3:12])
ggpairs(sound_hab[,3:13]) # added ACI & H
ggpairs(sound_hab, columns = c(3,4,5,6,7,8,9,10,11), 
        mapping = aes(color = as.factor(sound_hab$habitat)))

# COMMUNITY
sound_com = ecometrics %>%
  select(site, habitat, depth_m_avg, temp_avg, sal_avg, total_abund_sum, comm_richness_sum, comm_evenness_sum, 
         piscivore, avg_spl_low, avg_spl_high)

#sound_com$log_abund = log10(sound_com$total_abund_sum)
#plot(sound_com$total_abund_sum)

ggpairs(sound_com[,3:11])

## only piscivore was interesting
# sound_trophic = ecometrics %>%
#   select(site, habitat, carnivore, piscivore, planktivore, 
#          avg_spl_low, avg_spl_high)
# 
# ggpairs(sound_trophic[,3:7])


```

## Build mixed model to explore variation in SPL

```{r}
library(nlme)

# HIGH FREQUENCY - 
#didn't include avg_blade_length even though strong correlation Possibly explore seagrass alone

highMod1 = lm(avg_spl_high ~ totalcover + oyster + richness + evenness +  
                temp_avg + sal_avg + comm_evenness_sum + piscivore + habitat, data = ecometrics)
summary(highMod1)
anova(highMod1)

highMod2 = lm(avg_spl_high ~ totalcover + oyster +
                temp_avg + comm_evenness_sum + habitat, data = ecometrics)
summary(highMod2)
anova(highMod2)

highMod3 = lm(avg_spl_high ~ totalcover + oyster + temp_avg, data = ecometrics)
summary(highMod3)
anova(highMod3)

# LOW FREQUENCY

soundMod1 = lm(avg_spl_low ~ totalcover + oyster + evenness + habitat + deploy, data = ecometrics)
summary(soundMod1)
anova(soundMod1)



```

## Plots exploring pairwise correlation and variation among habitat types

```{r}


# add habitat label and filter to only seagrass

#Benthic metrics
ggplot(aes(x = richness, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = richness_comb, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = evenness , y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = totalcover, y = avg_spl_low, color = habitat), data = ecometrics)+ theme_bw()+
  geom_point(size = 3)+
  geom_smooth(method = "lm", se = FALSE)+
  scale_color_discrete(name = "Habitat type")+
  ylab("Average sound pressure level (db re 1 uPa")+ 
  xlab("Average percent cover")+
  theme(axis.text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        legend.title = element_text(size = 18), 
        legend.text = element_text(size = 16))

#Habitat Metrics
ggplot(aes(x = seagrass_comb, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = saltmarsh, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = saltmarsh, y = avg_spl_high, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = oyster, y = avg_spl_high, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = oyster, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

# Community metrics
ggplot(aes(x = total_abund_sum, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = comm_evenness_sum, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = comm_richness_sum, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = X1_Depth_m, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()


```
