---
title: "Sound_hab_spl_correlations"
author: "Becca Van Hoeck"
date: "3/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and data
```{r}
library(ggplot2)
library(GGally)
library(dplyr)
library(tidyr)

ecometrics = read.csv("data/clean_data/ecometrics.csv")
#ACI = read.csv("data/clean_data/avg_ACI1024.csv")
#H_entropy = read.csv("data/clean_data/avg_H1024.csv")
BI = read.csv("data/clean_data/bioacousticIndex.csv")

#ecometrics$ACI = ACI$avg_ACI1024
#ecometrics$entropy= H_entropy$avg_H1024
ecometrics$BI_low = BI$BI_low
ecometrics$BI_high = BI$BI_high

# make deployment a factor to include as a random effect
ecometrics$deploy = factor(ecometrics$deploy_num.x)

# correct habitat assignment for Seagrass C
ecometrics$habitat[ecometrics$site_code == "SC"] = "seagrass"
```

## Explore correlations:
1. Examine correlation within habitat/community metrics and remove redundant variables (corr > 0.7)
2. Explore correlation between habitat/community metrics and Sound Pressure Level

Start with habitat then community to reduce the number of correlations viewing at a single time
```{r}
# HABITAT
# sound_hab = ecometrics %>%
#   select(site, habitat, totalcover, seagrass_comb, oyster, saltmarsh, richness, evenness,
#          avg_shoot_density, avg_blade_length,
#         avg_spl_low, avg_spl_high, BI_low, BI_high)

sound_hab = ecometrics %>%
  select(site, habitat, totalcover, oyster, saltmarsh, evenness,
         avg_shoot_density, avg_blade_length,
        avg_spl_low, avg_spl_high, BI_low, BI_high)

ggpairs(sound_hab[,3:12])
#ggpairs(sound_hab[,3:13]) # added ACI & H
#ggpairs(sound_hab, columns = c(3,4,5,6,7,8,9,10,11), 
#        mapping = aes(color = as.factor(sound_hab$habitat)))

# COMMUNITY
sound_com = ecometrics %>%
  select(site, habitat, depth_m_avg, temp_avg, sal_avg, total_abund_sum, comm_richness_sum, comm_evenness_sum, 
         piscivore, avg_spl_low, avg_spl_high, BI_low, BI_high)

#sound_com$log_abund = log10(sound_com$total_abund_sum)
#plot(sound_com$total_abund_sum)

ggpairs(sound_com[,4:13])

## only piscivore was interesting
# sound_trophic = ecometrics %>%
#   select(site, habitat, carnivore, piscivore, planktivore, 
#          avg_spl_low, avg_spl_high)
# 
# ggpairs(sound_trophic[,3:7])


```

## Build model to explore variation in SPL

1. Using predictor variables that crossed a 0.3 correlation threshold, examine simple linear model with all candidate predictors
2. Evaluate model fit and evidence to retain or drop each predictor until a final model is reached

** Need to consider whether a mixed-effects model structure is more appropriate - random effect of deployment, site? should habitat type be a fixed or random effect?, should any interactions be included?**

Questions: 
1. Do I need to standardize or transform predictors?
2. How to evaluate if lm is sufficient, or should include mixed effects?
3. How to incorporate the variable effects of habitat? Can I exclude mudflats?

```{r}
#library(nlme)

# HIGH FREQUENCY - 
#didn't include avg_blade_length even though strong correlation Possibly explore seagrass alone

highMod1 = lm(avg_spl_high ~ totalcover + oyster + evenness + depth_m_avg + 
                temp_avg + sal_avg + comm_evenness_sum + piscivore + habitat, data = ecometrics)
summary(highMod1)
drop1(highMod1)

highMod2 = lm(avg_spl_high ~ totalcover + depth_m_avg + 
                temp_avg  + comm_evenness_sum + piscivore + habitat, data = ecometrics)
summary(highMod2)
drop1(highMod2)

highMod3 = lm(avg_spl_high ~ totalcover + depth_m_avg + temp_avg + habitat, data = ecometrics)
summary(highMod3)
drop1(highMod3)

highMod4 = lm(avg_spl_high ~ totalcover + depth_m_avg, data = ecometrics)
summary(highMod4)
drop1(highMod4)

AIC(highMod1, highMod2, highMod3, highMod4)
anova(highMod1, highMod2, test = "F") # how to interpret this?

## this doesn't work due to different sizes - missing data
#ecometrics$highmodpredict = predict(highMod2)

# Selected highMod2 due to lowest ACI, but the conf intervals for many of the effects overlap with zero
# doesn't this suggest no effect of those variables?
confint(highMod2)

# inspect residuals
residplot <- function(model){
  ggplot(model,aes(x=.fitted,y=.resid))+ 
    geom_point()+
    labs(x="Fitted values",y="Residuals")
  }
# inspect quantiles 
qqplot <- function(model){
  q1 <- ggplot(model,aes(sample=.stdresid))
  q1+geom_point(stat="qq")+geom_abline(slope=1,intercept=0)+
    labs(x="Theoretical Quantiles",y="Standardized residuals")
}

residplot(highMod2)
qqplot(highMod2)
confint(highMod2)

# LOW FREQUENCY 
#should I include interaction between habitat and total cover?

lowMod1 = lm(avg_spl_low ~ oyster + depth_m_avg + total_abund_sum + comm_richness_sum +
                 comm_evenness_sum + habitat, data = ecometrics)
summary(lowMod1)
drop1(lowMod1)

lowMod2 = lm(avg_spl_low ~ oyster + depth_m_avg + 
                 comm_evenness_sum, data = ecometrics)
summary(lowMod2)
drop1(lowMod2)

AIC(lowMod1, lowMod2)

# Selected lowMod2 for trial, but really not distinguishable by AIC
confint(lowMod2)

residplot(lowMod2)
qqplot(lowMod2)

```

## Variation within habitat type

```{r}
# Seagrass
sgMetrics = ecometrics %>%
  filter(habitat == "seagrass")

sound_seagrass = sgMetrics %>%
  select(site, habitat, totalcover, seagrass_comb, oyster, saltmarsh, richness, evenness,
         avg_shoot_density, avg_blade_length,
        avg_spl_low, avg_spl_high, BI_low, BI_high)

ggpairs(sound_seagrass[,3:14])

# # Creek
# creekMetrics = ecometrics %>%
#   filter(habitat == "creek")
# 
# sound_creek = creekMetrics %>%
#   select(site, habitat, totalcover, seagrass_comb, oyster, saltmarsh, richness, evenness,
#          avg_shoot_density, avg_blade_length,
#         avg_spl_low, avg_spl_high, BI_low, BI_high)
# 
# ggpairs(sound_creek,3:14)



```

## Plots exploring pairwise correlation and variation among habitat types

```{r}

# add habitat label and filter to only seagrass

#Benthic metrics
ggplot(aes(x = richness, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = richness_comb, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = evenness , y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = totalcover, y = avg_spl_low, color = habitat), data = ecometrics)+ theme_bw()+
  geom_point(size = 3)+
  geom_smooth(method = "lm", se = FALSE)+
  scale_color_discrete(name = "Habitat type")+
  ylab("Average sound pressure level (db re 1 uPa")+ 
  xlab("Average percent cover")+
  theme(axis.text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        legend.title = element_text(size = 18), 
        legend.text = element_text(size = 16))

#Habitat Metrics
ggplot(aes(x = seagrass_comb, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = saltmarsh, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = saltmarsh, y = avg_spl_high, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = oyster, y = avg_spl_high, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = oyster, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

# Community metrics
ggplot(aes(x = total_abund_sum, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = comm_evenness_sum, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = comm_richness_sum, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()

ggplot(aes(x = X1_Depth_m, y = avg_spl_low, color = habitat), data = ecometrics)+
  geom_point()


```
